Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-12-09 patchy-selection-paper.tex
added section infront of each section ref

@@

These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
To be clear about the assumptions (especially regarding scaling) implicit here,
we provide a derivation below in \nameref{apx:eqfreq}, 
as well as justification for the asymptotics below in \nameref{apx:asymptotics}.
In one dimension, the equation can be solved to give $q(x)$ 
in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.

@@

The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
by the ratio of the second terms in the two equations,
i.e.
\begin{align} \label{eqn:ratio_gestalt}
  \migrate(x) 
    &= \rho q(x) \times 
    \frac{
        \text{ ( probability a family establishes in patch at $x$ ) }
    }{
        \text{ ( mean occupation density of a family at $x$ ) }
    } .
\end{align}
To make this precise we need to understand what we mean by ``outflux of families'';
we do this below in \nameref{ss:hitting_occupation},
and provide a heuristic argument here.

@@

Suppose the family arrives with $K$ individuals,
i.e.\ there are $K$ individuals that manage to establish
The probability that at least one individual leaves a lineage that establishes in the patch
is $1-(1-p_e)^K$.
On the other hand, in the absence of the patch of new habitat,
each of the $K$ newly arrived individuals would leave behind a lineage with 
mean total size $\int_0^\infty e^{-t s_m} dt = 1/s_m$,
roughly half of which would stay in the new patch.
Therefore, the expected contribution of the family to the occupation density is $K/2s_m$.
(Readers who feel uncomfortable about the geometry of the patch,
or the distinction between occupation density and number of individuals
are encouraged to take a detour through \nameref{ss:hitting_occupation}, below.)

@@

To test the robustness of our results to the various approximations we used,
we used individual-based simulations on a one-dimensional lattice of 501 demes,
run for 25,000 generations. \gc{We simulated $\rho$ diploid
  individuals per patch, with nearest neighbor migration and
  multiplicative selection??!? }
More details are given below in \nameref{ss:simulations},
and the number of simulations used and parameter values are given in
supplementary tables \ref{stab:mutation_params} and
\ref{stab:migration_params}. 
To estimate the mean time until adaptation by mutation,
we used one centrally located patch of 99 demes
and a mutation rate of $\mu=10^{-5}$,
and to estimate the mean time until adaptation by migration,
we used two centrally located patches of 99 demes
separated by a varying number of demes,
with one patch initialized to have the $B$ allele at frequency 0.8.
In each case, we measured the ``time to adaptation'' as the amount of time until at least 100 $B$ alleles
were present in the previously unadapted patch.
Figure~\ref{fig:sim_times}
summarizes how the results compare to theory,
excluding parameter combinations where a majority of simulations did not adapt by 25,000 generations
or the patch was not large enough to support a stable population of $B$ alleles.
Supplementary figures \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} 
show all times, 
and depictions of typical simulation runs are shown in supplementary figures
\ref{sfig:sims_1}, \ref{sfig:sims_2}, \ref{sfig:sims_3}, \ref{sfig:sims_4}, \ref{sfig:sims_5}, \ref{sfig:sims_6}, \ref{sfig:sims_7}, and \ref{sfig:sims_8}.

@@

\paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
To provide context for the results on shared haplotype length in \nameref{ss:haplotype_length}
it is important to also understand
the process by which haplotypes are whittled down within patches.
The initial haplotype that sweeps within a patch 
will be dispersed over time by recombination.
Likewise, the haplotype that is shared between patches coadapted by migration
(equation~\ref{eqn:haplotype_length}) will also break down. 
However, a long time after the initial sweep,
we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
than with individuals elsewhere,
since selection against migrants 
decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
but more work is needed
to provide robust estimates for these processes.
\gc{We note that questions about the genomic length-scale of signals of
sweeps shared by migration have also been addressed in discrete
population settings \citep{REFSCOMMENTEDBELOW}. This work has shown
that the length of shared sweept haplotype is often significantly
shorter than the sweep with each patch, resulting in a pattern of
shoulders ofelevated $F_{ST}$ between adapted populations some distance away from
the shared selected allele. It would be of interest to see how similar
patterns can arise in a continuous population setting, as a way of
uniting these results. }

@@

Here we make a more precise argument to back up expression \eqref{eqn:migrate} for the migration rate.
The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
since it relies only on a decomposition of the migrant families upon hitting the new patch;
but it is also imprecise in a number of subtle ways
that are difficult to formalize.
Here we take a somewhat different tack,
supposing that it suffices to model the spatial movement of a migrant family
by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}),
and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
(Recall $\sigma$ is the dispersal distance.)
We then use facts about Brownian motion and branching processes, 
to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
\linelabel{rr:moved_argument}

>>>>

\gc{GENERAL GRAHAM comment, every place where we give a section name nameref
I've made sure that we say section nameref as otherwise it looks weird  }

@@

These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
To be clear about the assumptions (especially regarding scaling) implicit here,
we provide a derivation below in section \nameref{apx:eqfreq}, 
as well as justification for the asymptotics below in section \nameref{apx:asymptotics}.
In one dimension, the equation can be solved to give $q(x)$ 
in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.

@@

The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
by the ratio of the second terms in the two equations,
i.e.
\begin{align} \label{eqn:ratio_gestalt}
  \migrate(x) 
    &= \rho q(x) \times 
    \frac{
        \text{ ( probability a family establishes in patch at $x$ ) }
    }{
        \text{ ( mean occupation density of a family at $x$ ) }
    } .
\end{align}
To make this precise we need to understand what we mean by ``outflux of families'';
we do this below in section \nameref{ss:hitting_occupation},
and provide a heuristic argument here.

@@

Suppose the family arrives with $K$ individuals,
i.e.\ there are $K$ individuals that manage to establish
The probability that at least one individual leaves a lineage that establishes in the patch
is $1-(1-p_e)^K$.
On the other hand, in the absence of the patch of new habitat,
each of the $K$ newly arrived individuals would leave behind a lineage with 
mean total size $\int_0^\infty e^{-t s_m} dt = 1/s_m$,
roughly half of which would stay in the new patch.
Therefore, the expected contribution of the family to the occupation density is $K/2s_m$.
(Readers who feel uncomfortable about the geometry of the patch,
or the distinction between occupation density and number of individuals
are encouraged to take a detour through section \nameref{ss:hitting_occupation}, below.)

@@

To test the robustness of our results to the various approximations we used,
we used individual-based simulations on a one-dimensional lattice of 501 demes,
run for 25,000 generations. \gc{We simulated $\rho$ diploid
  individuals per patch, with nearest neighbor migration and
  multiplicative selection??!? }
More details are given below in section \nameref{ss:simulations},
and the number of simulations used and parameter values are given in
supplementary tables \ref{stab:mutation_params} and
\ref{stab:migration_params}. 
To estimate the mean time until adaptation by mutation,
we used one centrally located patch of 99 demes
and a mutation rate of $\mu=10^{-5}$,
and to estimate the mean time until adaptation by migration,
we used two centrally located patches of 99 demes
separated by a varying number of demes,
with one patch initialized to have the $B$ allele at frequency 0.8.
In each case, we measured the ``time to adaptation'' as the amount of time until at least 100 $B$ alleles
were present in the previously unadapted patch.
Figure~\ref{fig:sim_times}
summarizes how the results compare to theory,
excluding parameter combinations where a majority of simulations did not adapt by 25,000 generations
or the patch was not large enough to support a stable population of $B$ alleles.
Supplementary figures \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} 
show all times, 
and depictions of typical simulation runs are shown in supplementary figures
\ref{sfig:sims_1}, \ref{sfig:sims_2}, \ref{sfig:sims_3}, \ref{sfig:sims_4}, \ref{sfig:sims_5}, \ref{sfig:sims_6}, \ref{sfig:sims_7}, and \ref{sfig:sims_8}.

@@

\paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
To provide context for the results on shared haplotype length in
section \nameref{ss:haplotype_length}
it is important to also understand
the process by which haplotypes are whittled down within patches.
The initial haplotype that sweeps within a patch 
will be dispersed over time by recombination.
Likewise, the haplotype that is shared between patches coadapted by migration
(equation~\ref{eqn:haplotype_length}) will also break down. 
However, a long time after the initial sweep,
we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
than with individuals elsewhere,
since selection against migrants 
decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
but more work is needed
to provide robust estimates for these processes.
\gc{We note that questions about the genomic length-scale of signals of
sweeps shared by migration have also been addressed in discrete
population settings \citep{REFSCOMMENTEDBELOW}. This work has shown
that the length of shared sweept haplotype is often significantly
shorter than the sweep with each patch, resulting in a pattern of
shoulders ofelevated $F_{ST}$ between adapted populations some distance away from
the shared selected allele. It would be of interest to see how similar
patterns can arise in a continuous population setting, as a way of
uniting these results. }

@@

Here we make a more precise argument to back up expression \eqref{eqn:migrate} for the migration rate.
The argument made above in section \nameref{ss:heuristics} applies to general migration mechanisms,
since it relies only on a decomposition of the migrant families upon hitting the new patch;
but it is also imprecise in a number of subtle ways
that are difficult to formalize.
Here we take a somewhat different tack,
supposing that it suffices to model the spatial movement of a migrant family
by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}),
and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
(Recall $\sigma$ is the dispersal distance.)
We then use facts about Brownian motion and branching processes, 
to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
\linelabel{rr:moved_argument}
